### 2024-05-09 paper/paper.tex
Text

%%
%% This is file `sample-sigplan.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,sigplan')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigplan.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmISBN{978-1-4503-XXXX-X/18/06}

%% get rid of this for the final version
\usepackage{todonotes}
\usepackage{listingsutf8}
\setlength {\marginparwidth }{2cm} %This is for todod
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{cryptocode}
\usepackage{lstcoq}
\usepackage[utf8]{inputenc}
%% End of get rid of this for the final version

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The Name of the Title Is Hope}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Mukesh Tiwari}
\affiliation{%
 \institution{Swansea University}
 \country{United Kingdom}}
 \email{mukesh.tiwari@swansea.ac.uk}

\author{Berry Schoenmakers}
\affiliation{%
  \institution{TU Eindhoven, Eindhoven}
  \country{The Netherlands}}
  \email{l.a.m.schoenmakers@tue.nl}

\author{Bas Spitters}
\affiliation{%
  \institution{Aarhus University}
  \country{Denmark}}
\email{spitters@cs.au.dk}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Tiwari et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Despite the recent growth of zero-knowledge succinct argument 
  of knowledge (ZKSNARK) in cryptocurrencies, 
  sigma protocol remains a popular choice in electronic voting to maintain privacy 
  and verifiability of elections and still used in prominent e-voting systems such as 
  Helios, Belenois, and SwissPost \textbf{More examples}. However, 
  the existing implementations of sigma protocol have two 
  shortcomings: (i) the same code is written twice, once for the front-end (JavaScript)
  and once for the back-end (Python/OCaml/Java), which is time-consuming but more importantly error-prone, 
  and (ii) none of these implementations offer any security guarantees. 
  In this paper, we present a certified implementation of sigma protocol in the Coq proof assistant. 
  We prove that our implementation is (special) sound, complete, and 
  (honest verifier) zero-knowledge. Moreover, 
  we compile our Coq formalisation to WebAssembly and Rust, and 
  thereby solve the problem of writing the same code twice. 
  We demonstrate the applicability of our
  formalisation by encoding the cryptographic protocols implemented in 
  Helios, Belenois, and SwissPost voting systems. 
  Our long terms goal is to provide a verified cryptographic library that can be 
  readily used in privacy-preserving voting systems.

\end{abstract}

\begin{CCSXML}
<ccs2012>
  <concept>
      <concept_id>10002978.10002986.10002990</concept_id>
      <concept_desc>Security and privacy~Logic and verification</concept_desc>
      <concept_significance>500</concept_significance>
      </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Security and privacy~Logic and verification}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Sigma Protocol, Formal Verification, Coq, Cryptography, Zero-Knowledge Proof}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{How to pitch this paper and What is the unique selling point of our work?}
Our Coq formalisation can be used at front-end (WebAssembly code) and 
back-end (Rust). Contrary to the existing work, our formalisation
does not write the same code twice. But sigma protocol is arcane and has 
been formalised couple of times before so why another formalisation?
None of the formalisation has been able to extract the code to executable code 
and one that was able to do so did not reason about probabilities in an intuitive way.
Moreover, that work was just geared towards verifying a transcript rather than
constructing it. Also, recently ZKP community started a movement to 
formalise the sigma protocols. We also demonstrates the 
applicability of our formalisation by encoding Helios cryptographic protocols
in our formalisation.

\textbf{We need to explain why formal verification is important in the context of voting systems.}

\begin{itemize}
  \item Explain zero-knowledge proofs (completeness, soundness, zero-knowledge)
  \item Explain sigma protocols (completeness, special-soundness, honest-verifier zero-knowledg) 
  and mention about the Schnorr protocol
  \item Our contribution
  \item Explain the Schnorr protocol
  \item Explain Parallel, And, Eq, Or, and NEQ relations (also explain that 
  most of the formalisation has worked on And and Or relations)
\end{itemize}

\section{Introduction}

\textbf{Emphasise how Sigma protocol is still used in elections to a public office}
Despite the recent growth and popularity of zero-knowledge succinct 
argument of knowledge (ZKSNARK), sigma protocol remains a
leading cryptographic proof system in privacy-preserving voting systems. 
Sigma protocol is used in Helios --a widely popular voting system-- \cite{adida2008helios}, 
Belenois --used in French elections for overseas 
voters\footnote{VVFE, used in French elections, is a derivative of (part of) Belenios and 
inherited all the zero-knowledge proofs from Belenois.}-- \cite{cortier2023french}, and 
SwissPost --used in Swiss elections for overseas voters-- voting 
systems \cite{10.1007/978-3-031-15911-4_4} (\textbf{Look for more examples}). 
Moreover, recently zero-knowledge proof community started a movement to standardise
the sigma protocol \cite{ZKProof}.
One of the main reasons for its popularity in privacy-preserving voting systems is its simplicity and 
well-understood security.

%\textbf{Introduce here that sigma protocol is efficient and simple and all other details about sigma protocol.} 
Zero-knowledge proofs are possible for all problems in the $NP$ complexity class, which 
includes problems where a proof of membership can be efficiently verified in polynomial time.
Sigma protocol is a simple and efficient zero-knowledge proof where the verifier  is assumed to be honest. 
More formally, for a given $NP$-relation $R$ and a statement $x$, a sigma protocol allows a prover 
$P$ to convince a verifier $V$ that they posses a witness $w$ --polynomial in length of $x$-- such 
that ($x$, $w$) $\in$ $R$. To accomplish this, $P$ and $V$ engage in a three move 
interactive protocol (shown below). At the end of the protocol $V$ either accepts or rejects the proofs.
Like any other proof system, sigma protocols also has completness and (special) soundness; moreover,  
it also has an additional property (special honest-verifier) zero-knowledgeness.

\textbf{Draw a move }

\definition{Completeness}
when $P$ and $V$ follow the protocol, then $V$ always accepts. In other words, 
if $P$ knows $w$ for $x$ then it would successfully construct a triple 
(a; c; r) that always passes the verification equation, i.e., $V$ accepts.

\definition{Special soundness}
there exists a probabilistic polynomial time extractor $E$ which given 
two accepting transcripts (a; c; r) and (a; c', r') for $x$, then 
it can extract a witness $w$ such that $(x, w) \in R$. In other words, 
if $P$ does not know $w$ for $x$ then it cannot create a proof 
(a; c; r) that will passes the verification check.

\definition{Special honest-verifier zero-knowledge proof}
there exists a probabilistic polynomial-time simulator S that for  
any $v$ in $L_R$ and $c \in C$, $S$ generates conversations 
(a,c,r) with the same probability as honest conversations between 
$P$ and $V$ using $w$ satisfying $(v,w) \in R$.

Sigma protocol was first defined by Ronald Cramer \cite{cramer1996modular} 
and the first efficient sigma protocol was introduced by Schnorr in \cite{schnorr1991efficient}. 
Eventhough there exists many sigma protocols, the Schnorr protocol is one of the most well-known
sigma protocol and is used extensively in privacy-preserving voting systems, e.g, Helios, 
Belenois, SwissPost, etc. Therefore, our formalisation is geared towards the Schnorr protocol; moreover, we 
also formalise $n$ PARALLEL composition, $n$ AND composition, $n$ OR composition, 
and $n$ EQ composition of Schnorr protocol. We also formalise Okamoto protocol as a subprotocol 
for NEQ composition (\textbf{Berry: does it sound okay?}). However, we have not seen 
much use of Okamoto protocol in the privacy-preserving voting systems, 
at least in the existing voting systems. Ideally, given the abstraction of 
our formlisation (vector space) we could have abstracted the 
sigma protocol and their composition in a generic way \cite{10.1007/978-3-642-02384-2_17},
but we decided to focus on the Schnorr protocol and its composition to ensure nice and familiar APIs 
for e-voting in Web Assembly and Rust.

There are numerous implementations of sigma protocol in various languages, 
e.g., Helios is written in Python, Belenois is written in OCaml, and SwissPost is written in Java.
However, there are two main issues with these implementations: (i)  
the same code is written twice 
once for the front-end (JavaScript, TypeScript) and once for the back-end (Python, OCaml, or Java),
especially in privacy-preserving voting systems such as Helios, SwissPost, and 
(ii) none of these implementations offer any formal guarantees about the correctness of the code.
Given the importance of sigma protocols in privacy-preserving voting systems,
it has been formalised before  \cite{5552642,butler2019formalisation,10.1145/3319535.3354247}. 
\cite{5552642,butler2019formalisation} are mainly focused on establishing the 
pen-and-paper proofs of sigma protocol into theorem provers and not extracting an executable code.
On the other hand, \cite{10.1145/3319535.3354247} exatract an OCaml code from Coq 
formalisation but it is mostly geared towards verifying an Helios election transcript, 
rather than constructing the primitives to conduct an election. Moreover, 
it lacks many primitives and does not reason about probabilities in an intuitive way.

\subsection{Contribution}
In this paper, we present a certified implementation sigma protocol and its composition 
in the Coq proof assistant.

Moreover, we extract WebAssembly and Rust code from our 
certified implementation. 
We use our formalisation to model the cryptographic protocols implemented in Helios, 
Belenois, and SwissPost voting systems. To the best of our knowledge, this is the first
formal verification of Sigma protocol that can be used at the front-end (WebAssembly code) and
back-end (Rust code).

\section{Coq formalisation}

  We briefly explain the Schnorr protocol, which is very similar to 
  the one we explained in the introduction section.
  Given some public input $(G, g, q, h)$ where $G$
  is a cyclic group of prime order $q$, and $g$ and $h$ are two 
  generators of the group $G$, the prover claims that they know a
  witness $w$ for the statement $h = g^w$; the existence of such
  a $w$ is immediate because $g$ generates the group. However,
  does the prover know the witness $w$? In order to convince the
  verifier, the prover and the verifier do the following:
  \begin{itemize}
    \item the prover picks a random number $u$, computes $c = g^u$,
    and sends $c$ to the verifier.
    \item the verifier picks a random challenge $e$ and sends it to
    the prover
    \item the prover computes $t = u + e * w$ and sends $t$ to the
    verifier
  \end{itemize}

The verifier accepts if $g^t = c * h^e$, otherwise rejects. 
We model the Schnorr protocol in Coq as follows:

\begin{lstlisting}[language=Coq]
  Definition schnorr_protocol (x : F) (g : G) (u c : F) : 
    @sigma_proto 1 1 1 :=  ([g^u]; [c]; [u + c * x]).
\end{lstlisting}

 where $F$ is the underlying field, $G$ is the underlying group,
 $+ : F \rightarrow F \rightarrow F$ is field addition,
 $* : F \rightarrow F \rightarrow F$ is field multiplication, and
 $\textasciicircum : G \rightarrow F \rightarrow G$ scalor multiplication. 
 We emphasise that we work in abstract setting of \textbf{vector space} to 
 make the implementation as general as possible. Recall that a vector space consists of a set of 
 vectors, denoted by \(V\), a field \(F\) (elements of F are called scalars), and 
 two operations: vector addition and scalar multiplication.

\begin{itemize}
    \item \textit{Vector addition}: for any vectors \(\mathbf{v}, \mathbf{w} \in V\), their sum, denoted by \(\mathbf{v} + \mathbf{w}\), is also in \(V\).
    \item \textit{Scalar multiplication}: for any vector \(\mathbf{v} \in V\) and any scalar \(c \in F\), their product, denoted by \(c\mathbf{v}\), is also in \(V\).
\end{itemize}

These operations must satisfy the following properties for all vectors \(\mathbf{u}, \mathbf{v}, \mathbf{w} \in V\) and all scalars \(c, d \in F\):
(i) \textbf{Closure under addition:} \(\mathbf{u} + \mathbf{v} \in V\), (ii) \textbf{Commutativity of addition:} \(\mathbf{u} + \mathbf{v} = \mathbf{v} + \mathbf{u}\), 
(iii)  \textbf{Associativity of addition:} \((\mathbf{u} + \mathbf{v}) + \mathbf{w} = \mathbf{u} + (\mathbf{v} + \mathbf{w})\), 
(iv) \textbf{Existence of zero vector:} There exists a vector \(\mathbf{0} \in V\) such that \(\mathbf{v} + \mathbf{0} = \mathbf{v}\) for all \(\mathbf{v} \in V\), 
(v) \textbf{Existence of additive inverse:} For each vector \(\mathbf{v} \in V\), there exists a vector \(-\mathbf{v} \in V\) such that \(\mathbf{v} + (-\mathbf{v}) = \mathbf{0}\),
(vi) \textbf{Closure under scalar multiplication:} \(c\mathbf{v} \in V\), 
(vii) \textbf{Distributive properties:} \(c(\mathbf{u} + \mathbf{v}) = c\mathbf{u} + c\mathbf{v}\) and \((c + d)\mathbf{v} = c\mathbf{v} + d\mathbf{v}\), 
and (viii) \textbf{Compatibility with field multiplication:} \(c(d\mathbf{v}) = (cd)\mathbf{v}\) and \(1\mathbf{v} = \mathbf{v}\), where \(1\) denotes the multiplicative identity in \(F\).
In our setting we instantiate the set of vectors with group $G$ and field with $F$. Moreover, our notation for scalar multiplication 
is $\textasciicircum$ because it represents, when instantiated concretely, exponentiation function. Finally, 
we define two more functions

\begin{lstlisting}[language=Coq]
(* Fake transcript (without the witness x) *)
Definition schnorr_simulator (g h : G) (u c : F) : 
@sigma_proto 1 1 1 := ([gop (g^u) (h^(opp c))]; [c]; [u]).

(* 
  This function checks if a conversation (a; c; r) 
  is accepting or not. It checks if g^r = a * h^c
*)
Definition accepting_conversation 
  (g h : G) (v : @sigma_proto 1 1 1) : bool :=
  match v with
  | (a; c; r) =>  
    match Gdec (g^(hd r)) (gop (hd a) (h^(hd c))) with 
    | left _ => true
    | right _ => false 
    end
  end.

(* Distribution that involves the secret x *)
Definition schnorr_distribution  (lf : list F) 
  (Hlfn : lf <> List.nil) (x : F) (g : G) (c : F) : 
  dist sigma_proto :=
  (* draw u from a random distribution *)
  u <- (uniform_with_replacement lf Hlfn) ;;
  Ret (schnorr_protocol x g u c).

(* without secret x *)
Definition simulator_distribution (lf : list F) 
  (Hlfn : lf <> List.nil) (g h : G) (c : F) : 
  dist sigma_proto :=
  (* draw u from a random distribution *)
  u <- (uniform_with_replacement lf Hlfn) ;;
  Ret (schnorr_simulator g h u c).
\end{lstlisting}

Finally we establish the completeness, special soundness, 
and special honest-verifier zero-knowledge proof. 
\begin{lstlisting}[language=Coq]
Lemma schnorr_completeness : 
  forall (r c : F) (a : t G 1) (c r : t F 1),
  (a; c; r) = (schnorr_protocol x g r c) ->
  accepting_conversation g h (a; c; r) = true.
Proof. (* proof terms omitted *) Qed.

Lemma simulator_completeness : 
    forall (r c : F) (a : t G 1) (c r : t F 1),
    (a; c; r) = (schnorr_simulator g h r c) ->
    accepting_conversation g h (a; c; r) = true.
  Proof using -(x R). (* proof terms omitted *) Qed.

Lemma special_soundness: 
  forall (a : G) (ca ra cb rb : F), ca <> cb ->
  accepting_conversation g h ([a]; [ca]; [ra]) = true ->  
  accepting_conversation g h ([a]; [cb]; [rb]) = true ->
  (* then we can find a witness y such that g^y = h *)
  exists y : F, g^y = h /\ y = ((ra - rb) * inv (ca - cb)).
Proof using -(x R). (* proof terms omitted *) Qed.

(* it's identical, i.e., information theoretic zero-knowledge proof *)
(* Under the hood, it is modelled as a list and looks like:
    [((a; c; r), prob); ((a; c; r), prob) ......].
  We map accepting_conversation to crunch the first pair, 
  (a, c, r), and produce boolean a value (true), 
  and then we show that these two distribution are 
  identical 
*)
Lemma special_honest_verifier_zkp : 
  forall (lf : list F) (Hlfn : lf <> List.nil) (c : F), 
  List.map (fun '(a, p) => (accepting_conversation g h a, p))
    (@schnorr_distribution lf Hlfn x g c) = 
  List.map (fun '(a, p) => (accepting_conversation g h a, p))
    (@simulator_distribution lf Hlfn g h c).
Proof. (* proof terms omitted *) Qed.

\end{lstlisting}

In our library, we have encoded Parallel composition, 
AND composition, OR composition, EQ composition, and 
NEQ composition. Below we explain the OR composition
beause it is more involved than the rest.

  However, we would like 
  to emphasise that we worked in a vector-space, rather 
  than concrete numbers. Working in (abstract) vector-space makes our system 
  amenable to be instantiated by any group and field 
  (Elliptic curve vs Concrete Nubmers). \textbf{Currently, 
  our library comes with concrete nubmers but it can easily be 
  instantiated with Elliptic curve cryptography}

  \textbf{Coq code of Schnorr goes here}

  The reader can see that we have used vector space. Now 
  we establish completeness, (special) soundness, and zero-knowledge 
  proof of knowledge (ZkPOK).

  \textbf{Coq code soundness, completeness, zkp goes here}

  The proof of completeness and soundness are fairly straightforward but 
  zero-knowledge proof of knowledge requires probabilistic reasoning.

  \textbf{Coq code goes here}

  In our library, we have encoded Parallel composition, 
  And composition, Or composition, EQ composition, and 
  NEQ composition. Below we explain the OR composition
  beause it is more involved than the rest.

  \textbf{OR composition and Coq code}

  \subsection{Parallel Composition}
  Explain the parallel composition of sigma protocols.

  \subsection{And Composition}
  Explain the and composition of sigma protocols.

  \subsection{Or Composition}
  Explain the or composition of sigma protocols.

  \subsection{Equality Composition}
  Explain the equality composition of sigma protocols.

  \subsection{Inequality Composition}
  Explain the inequality composition of sigma protocols.

  \subsection{Uniform Distribution}
  To reason about probabilistic distribution, 
  we have encoded probability as list.

  \subsection{Fiat-Shamir Transform}
 We have formalised the interactive version of sigma protocol. However,
 interaction is not convenient and therefore
 we use Fiat-Shamir transform to make it non-interactive. We use SHA256 
 to turn the interaction to non-interaction but in order to 
 make it a self-contained library and 
 avoid relying on any external library at the front-end (WebAssembly)
 we have encoded the SHA256 in the Coq theorem prover, with usual correctness 
 properties.

 \subsection{Efficient Modular Arithmetic}
  In order to call the functions from our formalisation, 
  a user need to instantiate

\section{Web Assembly and Rust from the Coq Formalisation}
  Bas Spitter

\section{Case Studies/Experiment}
  We use our formalisation to model the Cryptographic protocols implemented in 
  Helios voting systems. Compare it with state-of-the-art 
  crypto library.

\section{Related Work}
  There are three work. Giles Barthe does not extract to executable code, 
  CryptoHol uses law of excluded middle, and thomas haines work is more about 
  verifying a transcript rather than constructing it. Moreover, 
  it does not reason about probabilities in intuitive way.

\section{Conclusion, Future Work, and Limitations}
  In this paper, we present a certified implementation 
  that can be readily used by voting community and 
  we demonstrated the usefulness by encoding Helios, 
  Belenois, and SwissPost using our primitives. 
  Moreover, our formalisation extracts to WebAssembly and 
  Rust without duplicate efforts.

\section{DECRA}
The aim of this project is to investigate the applicability of 
zero-knowledge succinct argument of knowledge (ZKSNARK) 
to reduce the data breaches.

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{reference}

%%
%% If your work has an appendix, this is the place to put it.
%\appendix

\end{document}
\endinput
%%
%% End of file `sample-sigplan.tex'.